Data exploration, quality control and statistical analysis of ChIP-exo experiments

### 2016-05-03 rmarkdown/paper.tex
fixed exploratory analysis, comparison with chip seq and calculate of chip -seq qc metrics with chip-exo and chip-nexus data

We first compared various factors that could affect binding site
identification between ChIP-exo and ChIP-Seq data. In order to compare
distribution of signal and background between ChIP-exo and ChIP-Seq
data, we calculated ChIP tag counts across the genome by counting the
number of reads mapping to each of 150 non-overlapping
window after extending reads by 150 to their $3\prime$
end directions. ChIP tag counts in ChIP-exo data were linearly related
to ChIP tag counts in ChIP-Seq data for the regions with high ChIP tag
counts (Upper part of Figure \ref{fig:comp}A). This implies that
signals for potential binding sites are well reproducible between
ChIP-exo and ChIP-Seq data. On the other hand, there was clear
difference in the background distribution between them (lower part of
Figure \ref{fig:comp}A). Specifically, in ChIP-Seq data reads were
almost uniformly distributed over background (non-binding) regions and
the ChIP tag counts in there regions were significantly larger than
zero. In contrast, in ChIP-exo data, there was larger variation in
ChIP tag counts among background regions and ChIP tag counts were much
lower in these regions compared to ChIP-Seq data. There were also
large proportion of regions without any read in ChIP-exo data. These
results indicate that for ChIP-exo data a much smaller portion of the
genome is expected to be background.

We next evaluated the ``peak pair'' assumption from Rhee and Pugh,
2011 \cite{exo1}, i.e. a peak of reads in the forward strand is
usually paired with a peak of reads in the reverse strand that is
located in the other site of the binding site. Note that currently
available ChIP-exo data analysis methods, such as Wang et al., 2014
\cite{mace}, Madrigal 2015 \cite{cexor} and Bardet et al., 2013
\cite{peakzilla} rely on this assumption. In order to evaluate this
assumption, we reviewed the proportion of reads in the forward strand
in ChIP-exo peaks such as at least one binding site is predicted in
both ChIP-exo and ChIP-Seq data. We found that strands of reads were
much less balanced in ChIP-exo data than in ChIP-Seq data in these
regions with potential binding sites (Fig. \ref{fig:comp}B) and this
indicates that the peak pair assumption might not hold in real
ChIP-exo data.

We evaluated ChIP-exo data for CTCF factor from human genome
\cite{exo1} to investigate issues specific to eukaryotic genomes for
binding sites identification. Figures \ref{fig:comp}C and
\ref{fig:comp}D display the bin-level average read counts against
mappability and GC content. Each data point is obtained by averaging
the read counts across bins with the same mappability of GC -
content. In Figure \ref{fig:comp}C it is shown that the ChIP-exo tag
counts linearly increases with the mappability score and in Figure
\ref{fig:comp}D it is shown that for GC - content below 0.6, the mean
ChIP tag count increases and for GC - content greater than 0.6 it
shows a decreasing trend. Kuan et al., 2011 \cite{mosaics} studied the
presence of the mappability and GC - content biases in ChIP-Seq's
background. It is not surprising to see these biases also present in
ChIP-exo data, since ChIP-Seq and ChIP-exo signal seems to be linearly
correlated for enriched regions (Figure \ref{fig:comp}A). Rozowsky et
al., 2009 \cite{peakseq} and Valouev et al., 2008 \cite{quest} provide
in depth analysis of the mappability and GC - content biases for
ChIP-Seq respectively. Finally, these results indicate that binding
site identification for ChIP-exo data benefits from using methods that
take into account of apparent sequence biases such as mappability and
GC content.

@@

\begin{table}[h!]
  \centering
  \begin{tabu} to\linewidth{X[-1]|X[-1]|X[-1]|X[-1]|X[-1]|X[-1]|X[-1]}
    \firsthline
    \textbf{Condition} & \textbf{Treatment} & \textbf{Rep.} & \textbf{Depth} & \textbf{NSC} & \textbf{RSC} & 
    \textbf{PBC}\\
    \hline 
    Aerobic & No Rif. & 1 & 13,961,493 & 103.15 & 2.0193 & 0.1399 \\
    Aerobic & No Rif. & 2 & 14,810,838 & 162.70 & 1.7805 & 0.1633 \\
    Anaerobic & No Rif. & 1 &  16,108,774 & 153.51 & 1.8035 & 0.1353 \\
    Anaerobic & No Rif. & 2 &  13,636,541 & 172.59 & 2.014 & 0.1532 \\ 
    Aerobic & No Rif. & 1 & 902,921 & 13.77 & 1.1270 & 0.2689\\
    Aerobic & No Rif. & 1 & 1,852,124 & 17.91 & 1.5275 & 0.2590\\
    Aerobic & Rif. 20 min & 2 & 2,104,427 & 29.60 & 1.2844 & 0.2584\\
    Aerobic & Rif. 20 min & 2 & 11,548,572 & 13.08 & 1.5122 & 0.1510 \\
    \lasthline
    \end{tabu}
    \caption{Current QC metrics applied to generated \emph{E. Coli} $\sig$ samples. NSC stands for Normalized 
             Strand Cross-Correlation, RSC stands for Relative Strand Cross-Correlation and PBC stands for 
             PCR Bottleneck Coefficient.}
\label{tab:qc_sig}
\end{table}

\begin{table}[h!]
  \centering
  \begin{tabu} to\linewidth{X[1.3]|X[2.2]|X[-1]|X[-1]|X[-1]|X[1.8]|X[-1]|X[-1]|X[-1]}
    \firsthline
    \textbf{Protocol} & \textbf{Organism} & \textbf{TF} & \textbf{Cell line} & \textbf{Rep.} &
    \textbf{Depth} & \textbf{NSC} & \textbf{RSC} &   \textbf{PBC} \\ 
    \hline
    \multirow{13}{4em}{ChIP-exo} & Human & CTCF & HeLa & & 48,478,450 & 16.02 & 1.1960 & 0.4579 \\
    & \multirow{3}{4em}{Human} & \multirow{3}{4em}{ER} & \multirow{3}{4em}{MCF-7} & 1 & 9,289,835 & 
    19.87 & 1.0127 & 0.8082 \\
    &  &  & & 2 & 11,041,833 & 21.48& 1.0063 & 0.8024\\
    &  &  & & 3 & 12,464,836 & 18.72& 1.0100 & 0.8203 \\    
    &  \multirow{3}{4em}{Mouse} & \multirow{3}{4em}{FoxA1} & \multirow{3}{4em}{Liver} & 1 & 22,210,461 & 21.28 & 1.1104 &  0.6562 \\
    &  & & & 2 & 23,307,557 & 60.42 & 1.1604 & 0.7996 \\ 
    &  & & & 3 & 22,421,72  & 72.04 & 1.1975 & 0.1068 \\
    & \multirow{3}{4em}{Human} & \multirow{3}{4em}{GR} & IM90 & & 47,443,803 & 8.86  & 1.3678 & 0.2978 \\
    & &  & K562 & & 116,518,000   & 4.11 & 1.0441 &0.0504 \\
    & &  & U2OS & & 3,255,111 &  10.05 & 1.0288 & 0.7714 \\
    & \multirow{3}{4em}{Human} & \multirow{3}{4em}{TBP} & \multirow{3}{4em}{K562} &1 & 61,046,382 &  12.01  & 1.1119
 & 0.1232 \\
    & & & &2 & 94,314,770 & 7.93 & 1.0299 & 0.1681 \\
    & & & & 3 & 114,282,270 & 9.25 & 1.1027 & 0.1464\\
\hline
\multirow{10}{4em}{ChIP-nexus} & \multirow{8}{4em}{D.Melanogaster} &
       \multirow{2}{4em}{Dorsal} & \multirow{4}{4em}{embryo} & 1 &   8,863,170 &  7.27 & 1.0402 &  0.6766 \\
 & & &  & 2 & 10,003,562 & 7.19 & 1.0672 & 0.5656\\
 & &  \multirow{2}{4em}{Twist} & & 1 & 18,244,203 & 5.82 &  1.1637 &  0.6592 \\
 & & &  & 2 & 52,546,982 & 5.27  &  1.1805  & 0.4549 \\
 & & \multirow{2}{4em}{Max} & \multirow{4}{4em}{S2} & 1 & 18,320,743 & 3.60 & 1.3628 & 0.5178  \\
 & & & & 2 & 24,965,642  & 3.47 & 1.0138 & 0.2124\\
 & & \multirow{2}{4em}{MyC} & & 1 & 7,832,034 & 5.92 & 1.0115 &  0.3935  \\
 & & & &  2 & 22,824,467 & 5.76 & 1.0045 & 0.1879\\
 & \multirow{2}{4em}{Human} & \multirow{2}{4em}{TBP} & \multirow{2}{4em}{K562} & 1 &  
              33,708,245 & 32.16 & 1.1712 & 0.3102 \\
 & &  &  & 2 & 129,675,001 & 32.70 &  1.2455 & 0.0492 \\
    \lasthline
  \end{tabu}
  \caption{Current QC metrics applied to gathered data. NSC stands for Normalized 
    Strand Cross-Correlation, RSC stands for Relative Strand Cross-Correlation and PBC stands for 
    PCR Bottleneck Coefficient.}  
\label{tab:qc}
\end{table}

DNA libraries generated by ChIP-exo and ChIP-nexus protocols are
expected to be less complex than the libraries generated by ChIP-Seq,
since the possible number of positions to which the reads can be
aligned is being reduced due to the exonuclease digestion,
considerable amounts of reads are being mapped to specific
positions. This affects the interpretation of the PBC, since for
ChIP-Seq low PBC values indicate that the same read has been copied by
the amplification process and aligned multiple times to the same
position; while for ChIP-exo when several reads are aligned to the
same position are not necessarily the same read amplified, but several
reads that their $5\prime$ end was digested to the same position
before the amplification step. It is of special importance to notice
that for deeply sequenced ChIP-exo and ChIP-nexus experiments, the PBC
values are quite low, which by following the ChIP-Seq guidelines it
would indicate that those experiment show severe bottlenecking
problems, which may incorrectly suggest that there is not genomic
signal int those experiments.

The Strand Cross-Correlation (SCC) introduced by Kharchenko et al.,
2008 \cite{strandcc} is the most commonly used quality measure in
ChIP-Seq. It is calculated as the correlation between both strand
coverages, where each one is shifted $\delta / 2$ bp towards the
$3\prime$ direction. In general it measures how well the reads mapped
to each strand are clustered around the locations where the proteins
are binding to the DNA, and usually it is expected to observed two
local maxima, one when the profiles are shifted by the average read
length and another when the profiles are shifted by the unobserved
fragment length. In a good ChIP-Seq dataset the last one is also the
SCC global maxima. However, in ChIP-exo's case these two peaks are
confounded. Hence the Normalized Strand Cross-Correlation (NSC) which
is a measure based on the SCC is harder to interpret. Figure
\ref{fig:scc_exo} shows that both local maxima are hard to
differentiate in the SCC curves for the $\sig$ ChIP-exo datasets used
to calculate the NSC values in Table \ref{tab:qc}.

@@

>>>>

We first compared various factors that could affect binding site
identification between ChIP-exo and ChIP-Seq data by using the
ChIP-exo reads from the first biological sample and first replicate
grown under aerobic condition (first line from Table \ref{tab:qc_sig})
and a SE ChIP-Seq replicate grown under the same conditions. In order
to compare distribution of signal and background between ChIP-exo and
ChIP-Seq data, we counted the number of extended reads mapping to a
partition of the genome into non-overlapping bins. ChIP tag counts in
ChIP-exo data were linearly related to ChIP tag counts in ChIP-Seq
data for the regions with high ChIP tag counts (Upper part of Figure
\ref{fig:comp}A). This implies that signals for potential binding
sites are well reproducible between ChIP-exo and ChIP-Seq data. On the
other hand, there was a clear difference in the background
distribution between them (lower part of Figure
\ref{fig:comp}A). Specifically, in ChIP-Seq data reads were almost
uniformly distributed over background (non-binding) regions and the
ChIP tag counts in there regions were significantly larger than
zero. In contrast, in ChIP-exo data, there was larger variation in
ChIP tag counts among background regions and ChIP tag counts were much
lower in these regions compared to ChIP-Seq data. There were also
large proportion of regions without any read in ChIP-exo data. These
results indicate that for ChIP-exo data a much smaller portion of the
genome is expected to be background. Therefore, methods that
specifically model the background distribution must take into account
that ChIP-exo's background is going to be composed by a smaller
proportion of the reads in the experiment.

Using the same experiment, we next evaluated the ``peak pair''
assumption from Rhee and Pugh, 2011\nocite{exo1}, i.e. a peak of reads
in the forward strand is usually paired with a peak of reads in the
reverse strand that is located in the other site of the binding
site. Note that currently available ChIP-exo data analysis methods,
such as Wang et al., 2014\nocite{mace}, Madrigal 2015\nocite{cexor}
and Bardet et al., 2013\nocite{peakzilla} rely on this assumption. In
order to evaluate this assumption, we reviewed the proportion of reads
in the forward strand in high quality ChIP-exo peaks such as at least
one binding site is predicted in both ChIP-exo and ChIP-Seq data. We
found that strands of reads were much less balanced in ChIP-exo data
than in ChIP-Seq data in these regions with potential binding sites
(Fig. \ref{fig:comp}B) and this indicates that the peak pair
assumption might not hold in ChIP-exo data. This is caused by either
the enzyme digestion or the strength at which the protein binds to the
DNA: Although it is expected for the exonuclease enzyme to digest the
ChIP fragments starting by their $5\prime$ ends and stopping when
finding a binding event, it may not be able to reach and digest every
fragment in the ChIP sample. On the other hand, if the protein is not
bound to both DNA strand, the enzyme may completely digest the
fragment. Finally both effect are increased by the PCR amplification.

We evaluated ChIP-exo data for CTCF factor from human genome
\cite{exo1} to investigate issues specific to eukaryotic genomes for
binding sites identification. Figures \ref{fig:comp}C and
\ref{fig:comp}D display the bin-level average read counts against
mappability and GC content. Each data point is obtained by averaging
the read counts across bins with the same mappability of GC -
content. In Figure \ref{fig:comp}C it is shown that the ChIP-exo tag
counts linearly increases with the mappability score and in Figure
\ref{fig:comp}D it is shown that for GC - content below 0.6, the mean
ChIP tag count increases and for GC - content greater than 0.6 it
shows a decreasing trend. Benjamini and Speed,
2011\nocite{benjamini2011} and Kuan et al., 2011\nocite{mosaics}
studied the presence of the mappability and GC - content biases in
ChIP-Seq's background. It is not surprising to see these biases also
present in ChIP-exo data, since ChIP-Seq and ChIP-exo signal seems to
be linearly correlated for enriched regions (Figure
\ref{fig:comp}A). Rozowsky et al., 2009\nocite{peakseq} and Valouev et
al., 2008\nocite{quest} provide in depth analysis of the mappability
and GC - content biases for ChIP-Seq respectively. Finally, these
results indicate that binding site identification for ChIP-exo data
benefits from using methods that take into account of apparent
sequence biases such as mappability and GC content, mostly when an
Input sample is not available.

@@

\begin{table}[h!]
  \centering
  \begin{tabu} to\linewidth{X[-1,m,c]|X[-1]|X[-1]|X[-1,m,c]|X[-1]|X[-1]|X[-1]|X[-1]}
    \firsthline
    \textbf{Bio Sample} & \textbf{Condition} & \textbf{Treatment} & \textbf{Rep.} & \textbf{Depth} & \textbf{NSC} & \textbf{RSC} & 
    \textbf{PBC}\\
    \hline 
    \multirow{4}{1em}{1} & Aerobic & No Rif. & 1 & 13,961,493 & 103.15 & 2.0193 & 0.1399 \\
    & Aerobic & No Rif. & 2 & 14,810,838 & 162.70 & 1.7805 & 0.1633 \\
    & Anaerobic & No Rif. & 1 &  16,108,774 & 153.51 & 1.8035 & 0.1353 \\
    & Anaerobic & No Rif. & 2 &  13,636,541 & 172.59 & 2.014 & 0.1532 \\ 
    \hline
    \multirow{4}{1em}{2} & Aerobic & No Rif. & 1 & 902,921 & 13.77 & 1.1270 & 0.2689\\
    & Aerobic & No Rif. & 1 & 1,852,124 & 17.91 & 1.5275 & 0.2590\\
    & Aerobic & Rif. 20 min & 2 & 2,104,427 & 29.60 & 1.2844 & 0.2584\\
    & Aerobic & Rif. 20 min & 2 & 11,548,572 & 13.08 & 1.5122 & 0.1510 \\
    \lasthline
    \end{tabu}
    \caption{Current QC metrics applied to generated \emph{E. Coli} $\sig$ samples. NSC stands for Normalized 
             Strand Cross-Correlation, RSC stands for Relative Strand Cross-Correlation and PBC stands for 
             PCR Bottleneck Coefficient.}
\label{tab:qc_sig}
\end{table}

\begin{table}[h!]
  \centering
  \begin{tabu} to\linewidth{X[1.3]|X[2.2]|X[-1]|X[1.3]|X[-1,m,c]|X[1.8]|X[-1]|X[-1]|X[-1]}
    \firsthline
    \textbf{Protocol} & \textbf{Organism} & \textbf{TF} & \textbf{Cell line} & \textbf{Rep.} &
    \textbf{Depth} & \textbf{NSC} & \textbf{RSC} &   \textbf{PBC} \\ 
    \hline
    \multirow{13}{1em}{ChIP-exo} & Human & CTCF & HeLa & & 48,478,450 & 16.02 & 1.1960 & 0.4579 \\
    & \multirow{3}{1em}{Human} & \multirow{3}{1em}{ER} & \multirow{3}{3em}{MCF-7} & 1 & 9,289,835 & 
    19.87 & 1.0127 & 0.8082 \\
    &  &  & & 2 & 11,041,833 & 21.48& 1.0063 & 0.8024\\
    &  &  & & 3 & 12,464,836 & 18.72& 1.0100 & 0.8203 \\    
    &  \multirow{3}{1em}{Mouse} & \multirow{3}{1em}{FoxA1} & \multirow{3}{1em}{Liver} & 1 & 22,210,461 & 21.28 & 1.1104 &  0.6562 \\
    &  & & & 2 & 23,307,557 & 60.42 & 1.1604 & 0.7996 \\ 
    &  & & & 3 & 22,421,72  & 72.04 & 1.1975 & 0.1068 \\
    & \multirow{3}{1em}{Human} & \multirow{3}{1em}{GR} & IMR90 & & 47,443,803 & 8.86  & 1.3678 & 0.2978 \\
    & &  & K562 & & 116,518,000   & 4.11 & 1.0441 &0.0504 \\
    & &  & U2OS & & 3,255,111 &  10.05 & 1.0288 & 0.7714 \\
    & \multirow{3}{1em}{Human} & \multirow{3}{1em}{TBP} & \multirow{3}{1em}{K562} &1 & 61,046,382 &  12.01  & 1.1119
 & 0.1232 \\
    & & & &2 & 94,314,770 & 7.93 & 1.0299 & 0.1681 \\
    & & & & 3 & 114,282,270 & 9.25 & 1.1027 & 0.1464\\
\hline
\multirow{10}{1em}{ChIP-nexus} & \multirow{8}{1em}{D.Melanogaster} &
       \multirow{2}{1em}{Dorsal} & \multirow{4}{1em}{embryo} & 1 &   8,863,170 &  7.27 & 1.0402 &  0.6766 \\
 & & &  & 2 & 10,003,562 & 7.19 & 1.0672 & 0.5656\\
 & &  \multirow{2}{1em}{Twist} & & 1 & 18,244,203 & 5.82 &  1.1637 &  0.6592 \\
 & & &  & 2 & 52,546,982 & 5.27  &  1.1805  & 0.4549 \\
 & & \multirow{2}{1em}{Max} & \multirow{4}{1em}{S2} & 1 & 18,320,743 & 3.60 & 1.3628 & 0.5178  \\
 & & & & 2 & 24,965,642  & 3.47 & 1.0138 & 0.2124\\
 & & \multirow{2}{1em}{MyC} & & 1 & 7,832,034 & 5.92 & 1.0115 &  0.3935  \\
 & & & &  2 & 22,824,467 & 5.76 & 1.0045 & 0.1879\\
 & \multirow{2}{1em}{Human} & \multirow{2}{1em}{TBP} & \multirow{2}{1em}{K562} & 1 &  
              33,708,245 & 32.16 & 1.1712 & 0.3102 \\
 & &  &  & 2 & 129,675,001 & 32.70 &  1.2455 & 0.0492 \\
    \lasthline
  \end{tabu}
  \caption{Current QC metrics applied to gathered data. NSC stands for Normalized 
    Strand Cross-Correlation, RSC stands for Relative Strand Cross-Correlation and PBC stands for 
    PCR Bottleneck Coefficient.}  
\label{tab:qc}
\end{table}

DNA libraries generated by ChIP-exo and ChIP-nexus protocols are
expected to be less complex than the libraries generated by ChIP-Seq,
since the possible number of positions to which the reads can be
aligned is being reduced due to the exonuclease digestion,
considerable amounts of reads are being mapped to specific
positions. This affects the interpretation of the PBC, since for
ChIP-Seq low PBC values indicate that the same read has been copied by
the amplification process and aligned multiple times to the same
position; while for ChIP-exo and ChIP-nexus when several reads are
aligned to the same position are not necessarily the same read
amplified, but several reads that their $5\prime$ end was digested to
the same position before the amplification step. It is of special
importance to notice that for deeply sequenced ChIP-exo and ChIP-nexus
experiments, the PBC values are quite low, which by following the
ChIP-Seq guidelines it would indicate that those experiment show
severe bottlenecking problems, which may incorrectly suggest that the
positions with a large amount of aligned reads are being caused by PCR
amplification rather than observed genomic signal.

The Strand Cross-Correlation (SCC) introduced by Kharchenko et al.,
2008\nocite{strandcc} is the most commonly used quality measure in
ChIP-Seq. In general it measures how well the reads mapped to each
strand are clustered around the locations where the proteins are
binding to the DNA, and usually it is expected to observed two local
maxima, one when the profiles are shifted by the average read length
and another when the profiles are shifted by the unobserved fragment
length. In a high quality ChIP-Seq dataset the last one is also the
SCC global maxima. On the other hand, as a thought experiment in an
idealized ChIP-exo experiment where the ChIP fragments are digested
until they found the binding protein and this protein bound to the
whole motif sequence, then the SCC would a flat function with a jump
at the motif length; since usually this is not the case, we also
expect to observe a local maxima at the read length. Figure
\ref{fig:scc_exo} shows the SCC curves for the CTCF factor from human
genome: The ChIP-exo curve shows a local maxima at the motif and read
lengths, while the SE ChIP-Seq curves have a local maxima at the read
length and a global maxima at the unobserved fragment
length. Therefore, in ChIP-exo's cases both peaks are likely to be
confounded. Hence measures based in the SCC such as the Normalized
Strand Cross-Correlation (NSC) or the Relative Strand
Cross-Correlation (RSC) are harder to interpret.

% that both
% local maxima are hard to differentiate in the SCC curves for the
% $\sig$ ChIP-exo datasets used to calculate the NSC values in Table
% \ref{tab:qc}.

% It is calculated as the correlation between both strand
% coverages, where each one is shifted $\delta / 2$ bp towards the
% $3\prime$ direction
\color{blue}

\begin{enumerate}
\item \st{Available data, the motivation of this part is two fold:}
 \begin{itemize}
 \item \st{we want to list all data sets we obtained}
 \item \st{we want to emphasize that for the sig70 datasets we have a
   unique design that allows us to compare chip exo with SE and PE
   chipseq , the later one for the first time}
 \end{itemize}
\item \st{comparison with chip seq} %? with and without dpeak ?}
\item \st{only in methods. list the align algorithm used}
\item \st{chip seq qc metrics}
\item chip exo qc pipeline, description and justification of the steps
\begin{itemize}
\item need to justify the sampling step. for this we may use the depth
  table and add the factor to normalize one depth to another. explain
  the confounding of pcr amplification and the reduction of the number
  of possible mappable positions.
\item case study with FoxA1, analysis of enrichment vs library
  complexity and analysis and the strand imbalance tools. in methods
  we need to add which tests did we perform to show that the pipeline
  works. how do we expect from the diagnostics and how to interpret
  the plots.
\item analysis of why the pipeline works. in methods need to specify
  how and in results need to specify why it works, and what are the
  conclusions from it
\end{itemize}
\item high resolution analysis with chip exo and comparison with PE
  chip seq
  \begin{itemize}
  \item justify what do we used dpeak (in methods need to list the
    other methods used). perhaps we can add gof plots in
    supplement. we also want to compare with the same method,
    therefore we can use the fact that dpeak have a unified model for
    PE and SE dpeak
  \item after comparing with other methods we can compare with PE and
    SE ChIP-seq, this comparison gives also resolution and sensitivity
  \item compare samples under same budget of reads
  \end{itemize} 
\end{enumerate}

@@

Using the $\sig$ sample grown under aerobic condition for ChIP-exo and
SE ChIP-Seq. We partitioned the \emph{E.Coli} genome into
non-overlaping bins of length 150, and for each bin we
counted how many ChIP-exo and SE ChIP-Seq extended reads overlapped
each bin. The reads were extended by 150 bp to their
$3\prime$ directions.

To model the strand imbalance, we called a set of high quality peaks
for ChIP-exo and SE ChIP-Seq (with $\mbox{GC content + Mappability}$
and $\mbox{Input only}$ models respectively), such that each peak
contained at least one binding event (found by using \textbf{dPeak}
with at most 5 binding events in each peak). For each peak, we
calculated the FSR as:

\begin{align*}
    \mbox{FSR} &= \frac{\text{Nr. of fwd. strand reads mapped to the region}}                 {\text{Total nr. of reads mapped to the region}} 
\end{align*}

We estimated the $\mbox{FSR}$'s densities by using R's \emph{density}
function using its default parameters, which computes gaussian kernel
density estimates.
